Cost- and Deadline-constrained Scheduling of Scientific Workflow Ensembles in IaaS Clouds

### 2012-01-20 paper/paper.tex
Ewa's changes

SPSS assigns sub-deadlines to each task based on the \emph{float time}
of the workflow. The float time is the amount of extra time that a workflow
can extend its critical path and still be completed by the ensemble deadline.
For a workflow $w$, the float time of $w$ is:

$$
FT(w) = d - CP(w)
$$

@@

SPSS distributes the float time of the workflow by level, so that each level
of the workflow gets a portion of the workflow's float time proportional to
the number of tasks in the level and the total runtime of tasks in the level.
The idea is that levels containing many tasks and large runtimes should be given
a larger portion of the float time so that tasks in those levels may be serialized.
Otherwise, many resources need to be allocated to run all of the tasks in 
parallel, which may be more costly.

The float time of a level $l$ in workflow $w$ is given by:

$$
FT(l) = FT(w) \Biggl[\left({\alpha}\frac{N(l)}{N(w)}\right) + \left({(1 - \alpha)}\frac{R(l)}{R(w)} \right)\Biggr]
$$

where $N(w)$ is the number of tasks in the workflow, $N(l)$ is the number of tasks 
in level $l$, $R(w)$ is the total runtime of all tasks in the workflow, $R(w)$ is
the total runtime of all tasks in level $l$, and $\alpha$ is a parameter between
0 and 1 that causes more float time to be given to levels with more tasks (large 
$\alpha$) or more runtime (small $\alpha$). We set $\alpha$ to be 0.7.

@@

$$
DL(t) = LST(t) + RT(t) + FT(Level(t))
$$

@@

$$
LST(t) = \begin{cases}
0,&$if t is a root task$\cr
\max_{p \in Pred(t)} DL(p),&$otherwise.$\cr
\end{cases}
$$

Algorithm~\ref{alg:planworkflow} shows how SPSS creates low-cost plans for
each workflow. The \textsc{PlanWorkflow} procedure first calls 
\textsc{DeadlineDistribution} to assign sub-deadlines to tasks according
to the procedure described above. Then, the \textsc{PlanWorkflow} procedure
schedules tasks onto VMs, allocating new VMs when necessary. For each task 
in the workflow, the the least expensive slot is chosen to schedule the 
task so that it can be completed by  its deadline. VMs are allocated in 
blocks of one billing cycle (one hour) regardless of the size of the task. 
When computing the cost of scheduling a task on a given VM, the algorithm 
considers idle slots in blocks that were allocated for previous tasks to be 
free, while slots in new blocks cost the full price of the new block. For 
example, if a task has a runtime of 10 minutes, and the price of a one-hour 
billing cycle is \$1, then the algorithm will either schedule the task on 
an existing VM that has an idle slot larger than 10 minutes for a cost of \$0, 
or it will allocate a new block on an existing VM, or provision a new VM, for 
a cost of \$1. If the cost of slots on two different VMs is equal, then the 
slot with the earliest start time is chosen. To prevent too many VMs from 
being provisioned, the algorithm always prefers to extend the runtime of 
existing VMs before allocating new VMs. The result of this is that the 
algorithm will only allocate a new block if there are no idle slots on 
existing blocks large enough or early enough to complete the 
task by its deadline, and it will only allocate a new VM if it cannot add 
a block to an existing VM to complete the task by its deadline.

@@

We show the results for three algorithms: DPDS, WA-DPDS and SPSS. These 
experiments were run with maximum autoscaling factor set to 1.0 for
DPDS and WA-DPDS. After experimenting with DPDS and WA-DPDS we found that, 
due to the high parallelism of workflows, the resource utilization remains 
high most of the time, which leads to the rapid provisioning of new resources 
at the beginning of ensemble execution. This, in turn, results in less 
efficient total utilization due to high levels of parallelism. Setting the 
scaling limit to 1.0, ensures that the shape of the resource provisioning 
plan  is close to a rectangle (see Fig.~\ref{fig:spds-example}) and the VM 
termination procedure (Algorithm~\ref{alg:prov}) allows maximum utilization 
of VMs before shutdown. For the SPSS algorithm we set the $\alpha$ parameter
for deadline distribution to be 0.7.

@@

We run the simulations for the constraint values ranging between the
estimated limits as indicated in Section~\ref{sec:scenarios}, resulting in
ca. 50-150 (budget, deadline) points for each application. The runs were
repeated for Pareto-distributed and constant workflow sizes. From this result
space we selected the most interesting plots which give the good overview  of
the observed algorithm behavior.

>>>>

SPSS assigns sub-deadlines to each task based on the slack time of the
workflow, which is defined as the amount of extra time that a workflow
can extend its critical path and still be completed by the ensemble deadline.
For a workflow $w$, the slack time of $w$ is:

$$
ST(w) = d - CP(w)
$$

@@

A task's level is the length of the longest path between the task and an entry
task of the workflow:

$$
Level(t) = \begin{cases}
0,&$if~$Pred(t) = \emptyset\cr
\max_{p \in Pred(t)}Level(p)+1,&$otherwise.$
\end{cases}
$$

SPSS distributes the slack time of the workflow by level, so that each level
of the workflow gets a portion of the workflow's slack time proportional to
the number of tasks in the level and the total runtime of tasks in the level.
The idea is that levels containing many tasks and large runtimes should be given
a larger portion of the slack time so that tasks in those levels may be serialized.
Otherwise, many resources need to be allocated to run all of the tasks in 
parallel, which may be more costly.

The slack time of a level $l$ in workflow $w$ is given by:

$$
ST(l) = ST(w) \Biggl[\left({\alpha}\frac{N(l)}{N(w)}\right) + \left({(1 - \alpha)}\frac{R(l)}{R(w)} \right)\Biggr]
$$

where $N(w)$ is the number of tasks in the workflow, $N(l)$ is the number of tasks 
in level $l$, $R(w)$ is the total runtime of all tasks in the workflow, $R(w)$ is
the total runtime of all tasks in level $l$, and $\alpha$ is a parameter between
0 and 1 that causes more slack time to be given to levels with more tasks (large 
$\alpha$) or more runtime (small $\alpha$).

@@

$$
DL(t) = LST(t) + RT(t) + ST(Level(t))
$$

@@

$$
LST(t) = \begin{cases}
0,&$if~$Pred(t) = \emptyset\cr
\max_{p \in Pred(t)} DL(p),&$otherwise.$
\end{cases}
$$

Algorithm~\ref{alg:planworkflow} shows how SPSS creates low-cost plans for
each workflow. The \textsc{PlanWorkflow} procedure first calls 
\textsc{DeadlineDistribution} to assign sub-deadlines to tasks according
to the procedure described above. Then, the \textsc{PlanWorkflow} procedure
schedules tasks onto VMs, allocating new VMs when necessary. For each task 
in the workflow, the the least expensive slot is chosen to schedule the 
task so that it can be completed by  its deadline. VMs are allocated in 
blocks of one billing cycle (one hour) regardless of the size of the task. 
When computing the cost of scheduling a task on a given VM, the algorithm 
considers idle slots in blocks that were allocated for previous tasks to be 
free, while slots in new blocks cost the full price of a billing cycle. For 
example, if a task has a runtime of 10 minutes, and the price of a block is 
\$1, then the algorithm will either schedule the task on 
an existing VM that has an idle slot larger than 10 minutes for a cost of \$0, 
or it will allocate a new block on an existing VM, or provision a new VM, for 
a cost of \$1. If the cost of slots on two different VMs is equal, then the 
slot with the earliest start time is chosen. To prevent too many VMs from 
being provisioned, the algorithm always prefers to extend the runtime of 
existing VMs before allocating new VMs. The result of this is that the 
algorithm will only allocate a new block if there are no idle slots on 
existing blocks large enough or early enough to complete the 
task by its deadline, and it will only allocate a new VM if it cannot add 
a block to an existing VM to complete the task by its deadline.

@@

We show the results for three algorithms: DPDS, WA-DPDS and SPSS. These 
experiments were run with maximum autoscaling factor set to 1.0 for
DPDS and WA-DPDS. After experimenting with DPDS and WA-DPDS we found that, 
due to the high parallelism of workflows, the resource utilization remains 
high most of the time, which leads to the rapid provisioning of new resources 
at the beginning of ensemble execution. This, in turn, results in less 
efficient total utilization due to high levels of parallelism. Setting the 
scaling limit to 1.0, ensures that the shape of the resource provisioning 
plan  is close to a rectangle (see Fig.~\ref{fig:spds-example}) and the VM 
termination procedure (Algorithm~\ref{alg:prov}) allows maximum utilization 
of VMs before shutdown. Based on experiments with different applications, 
we set the $\alpha$ parameter for deadline distribution in SPSS to be 0.7.

@@

We run the simulations for the constraint values ranging between the
estimated limits as indicated in Section~\ref{sec:scenarios}, resulting in
approximately 50-150 (budget, deadline) points for each application. The runs were
repeated for Pareto-distributed and constant workflow sizes. From this result
space we selected the most interesting plots which give the good overview  of
the observed algorithm behavior.
